---
title: "A Study of Beers and Breweries in the US"
author: "Ryan Paul & Chad Madding"
date: "October 18, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

*With a strong presence across the 50 states and the District of Columbia, craft breweries are a vibrant and flourishing economic force at the local, state and national level. As consumers continue to demand a wide range of high quality, full-flavored beers, small and independent craft brewers are meeting this growing demand with innovative offerings, creating high levels of economic value in the process.*

**Bart Watson**, Chief Economist for the Brewers Association

https://www.brewersassociation.org/statistics/economic-impact-data/

We have been hired to gain incite into breweries in the United Stated. We will focus on providing information to a national brewery for them to look into investing in the craft brewery industry.

### Data

The data for this study will consist of two databases collected and supplied by our client.

**Data/Beers.csv**

| Variable | Description |
| -------- | ----------- |
| Name | name of beer |
| BeerID | unique id number of the beer |
| ABV | alcohol by volume of beer |
| IBU | international bitterness units of the beer |
| Style | style of beer |
| Ounces | unit of measurement for the beer by oz |

**Data/Breweries.csv**

| Variable | Description |
| -------- | ----------- |
| Name | name of beer |
| Brew_ID | unique identification number of the brewery |
| Name | name of brewery |
| City | city where brewery is located |
| State | state where brewery is located |

### Folder Information

There are 3 folders in the BeersAndBreweries repository:

   1. Data - This folder contains the two databases provided by our client. (Beers.csv, Breweries.csv)
   2. Directives - This folder contains the original request for information. (Case Study 01.docx, CaseStudy1Rubric.docx)
   3. The CaseStudy01 folder simply contains the graphics used in this report.

R libraries used in the report
```{r libraries, message=FALSE, echo=TRUE}
#load libraries
library(dplyr)
library(plyr)
library(ggplot2)
library(pastecs)
```

Read the two databases into R
```{r read data, echo=TRUE}
#read in the two databases
breweries <- read.csv("Data/Breweries.csv")
beers <- read.csv("Data/Beers.csv")
```

### Questions our client would like addressed

   1. How many breweries are present in each state?
   
**The table below list breweries in each state. Washington DC, North Dakota, South Dakota and West Virginia are at the bottom with only one brewery. Colorado leads the list with 47.**

```{r breweries state, echo=TRUE}
#count breweries in each state and order from lowest to higest
brew_state <- sort(table(breweries$State))
#List out the table 
brew_state
```

   2. We will now merge the beer data with the breweries data and print the first 6 observations and the last six observations to check the new merged file.

```{r merge data, echo=TRUE}
#looking at each database to see problems before the merge
head(beers)
head(breweries)
#rename Brew_ID to Brewery_id to merge data
colnames(breweries)[1] <- "Brewery_id"
#both have a "Name" two for different groups
#rename Name to Brewery_Name
colnames(breweries)[2] <- "Brewery_Name"
#rename Name to Beer_Name
colnames(beers)[1] <- "Beer_Name"
#merge both data sets
brew_beer <- merge.data.frame(beers, breweries, by = "Brewery_id")
#print the first 6 observations
head(brew_beer, 6)
#print the last 6 observations
tail(brew_beer, 6)
```

   3. Report the number of NA's in each column.
   
**There are only two columns that have missing data. The ABV looks to be missing 62 entries and IBU has 1005 NA's out of the total 2410 observations. The IBU data may need to be revisited. With close to half the data not reported predictive information based on this information may be skewed. We will address this if this shows to be a problem.**
   
```{r count NAs, echo=TRUE}
#use colSum to sum up the NA's in each column
colSums(is.na(brew_beer))

```

   4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
   
```{r Compute the median, echo=TRUE}
#Compute the median alcohol content (ABV) by state
medABV <- tapply(brew_beer$ABV, brew_beer$State, function(x) median = median(x, na.rm = TRUE))
#Compute the median international bitterness unit (IBU) by state
medIBU <- tapply(brew_beer$IBU, brew_beer$State, function(x) median = median(x, na.rm = TRUE))

#plot a bar chart to compare ABV and IBU
#convert AVB and IBU to a data frame
medABVbar <- data.frame(medABV)
medIBUbar <- data.frame(medIBU)

#add a State column to ABV and IBU
medABVbar$State <- rownames(medABVbar)
medIBUbar$State <- rownames(medIBUbar)

#plot a bar chart for ABV
ggplot(na.omit(medABVbar), aes(x=reorder(medABVbar$State, medABVbar$medABV), y=medABVbar$medABV)) +
  geom_bar(stat="identity", color="white", fill="blue") +
  ylab("Alcohol Content") + xlab("State") +
  ggtitle("Median ABV by State") +
  coord_flip() + theme(plot.title = element_text(hjust=0.5), text = element_text(size=7.5))

#Remove SD from the list due to no IBU data
medIBUbar <- medIBUbar[-grep("SD", medIBUbar$State),]
#plot a bar chart for IBU
ggplot((medIBUbar), aes(x=reorder(medIBUbar$State, medIBUbar$medIBU), y=medIBUbar$medIBU)) +
  geom_bar(stat="identity", color="white", fill="blue") +
  ylab("IBU") + xlab("State") +
  ggtitle("Median IBU by State") +
  coord_flip() + theme(plot.title = element_text(hjust=0.5), text = element_text(size=7.5))

```

   5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

**In the dataset provided to us, Colorado has the higest alcoholic (ABV) beer and Oregon has the most bitter (IBU).**
      
```{r max ABV and IBU, echo= TRUE}
#state wuth the maximum alcoholic (ABV) beer
brew_beer$State[which.max(brew_beer$ABV)]
#state with the most bitter (IBU) beer
brew_beer$State[which.max(brew_beer$IBU)]
```

   6. Summary statistics for the ABV variable.

| Value | Description |
|-------|-------------|
| 2348 | the number of values (nbr.val) |
| 0 | number of null values (nbr.null) |
| 62 | number of missing values (nbr.na) |
| 0.001 | minimal value (min) |
| 0.128 | maximal value (max) |
| 0.127 | range (range, that is, max-min) |
| 140.348 | sum of all non-missing values (sum) |
| 0.056 | median (median) |
| 0.060 | mean (mean) |
| 0.00028 | standard error on the mean (SE.mean) |
| 0.00055 | confidence interval of the mean (CI.mean) at the p=0.95 level |
| 0.00018 | variance (var) |
| 0.01354 | standard deviation (std.dev) |
| 0.22655 | variation coefficient (coef.var) (sd/mean) |

```{r summary ABV, echo= TRUE}
#summary stats for ABV
options(scipen=100)
options(digits=2)
stat.desc(brew_beer$ABV)
```

   7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
   
```{r scatter plot, echo=TRUE}
#ggplot to check for correlation between IBU and ABV
ggplot(brew_beer, aes(x = IBU, y = ABV)) + geom_point(na.rm=TRUE)+geom_smooth(method=lm,se=FALSE, na.rm=TRUE)
```

**The scatter plot above looks to show a positive correlation between the ABV and IBU.**

**Looking at the numbers below we can see a very significant correlation between the alcohol content (ABV) and the bitterness (IBU) of a beer. One thing to keep in mind, there are 62 missing entries in AVB and 1005 in the IBU data leading to a lower degrees of freedom, but at a thousand in the current data set the correlation data is still relevant. This is observational data so even with the correlation any inferences made to the larger population would be speculative.**

```{r Pearson, echo+TRUE}
# Pearson to check for correlation 
cor.test(brew_beer$ABV, brew_beer$IBU, method = "pearson")
```

### Conclusion

Summarize your findings from this exercise.

**sessionInfo**

R version 3.5.1 (2018-07-02)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows >= 8 x64 (build 9200)
