---
title: "Brew Forward 2019"
author: "Chad Madding & Ryan Paul"
date: "10/16/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r librariesAndHelpers, message=FALSE}
#R libraries used in the report
#load libraries
list.of.packages <- c("plyr", "dplyr", "ggplot2", "pastecs", "reshape2", "kableExtra")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, repos = "http://cran.us.r-project.org")

library(plyr)
library(dplyr)
library(ggplot2)
library(pastecs)
library(reshape2)
library(kableExtra)
#function to trim white space
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

```{r readData}
#read in the two datasets
breweries <- read.csv("Data/Breweries.csv")
#States has a leading space, removing it
breweries$State <- trim(breweries$State)
beers <- read.csv("Data/Beers.csv")
#rename Brew_ID to Brewery_id to merge data
colnames(breweries)[1] <- "Brewery_id"
#both have a "Name" two for different groups
#rename Name to Brewery_Name
colnames(breweries)[2] <- "Brewery_Name"
#rename Name to Beer_Name
colnames(beers)[1] <- "Beer_Name"
#merge both data sets
brew_beer <- merge.data.frame(beers, breweries, by = "Brewery_id")
```

## The Campaign
With the beer and brewery data you have provided us, we have come up with the following observations:

- Craft breweries by state
- State specific beer property metrics
- Analysis of a potential relationship between IBU and ABV in beer

## State of Craft Breweries by State

```{r breweriesByState}
#Group breweries by state and provide a count
stateGroupedBreweryData <- breweries %>%
  group_by(State) %>%
  count()
#Changing column name of count
names(stateGroupedBreweryData)[2] <- "BreweriesCount"
#Creating a bar chart and ordering by states descending
stateBreweryCountBarChart <- ggplot(data=stateGroupedBreweryData,
                                    aes(x=reorder(State,-BreweriesCount), y=BreweriesCount, fill = State)) +
  geom_bar(stat="identity") +
  labs(y = "Craft Breweries", x = "State") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("Number of Craft Breweries by State") +
  guides(fill=FALSE) # Removing legend

#Display it
stateBreweryCountBarChart
```

## Notes on Missing Data
Some context regarding the upcoming analysis. We are missing the following values for the number of beers in the right column.

```{r countNAs}
#use colSum to sum up the NA's in each column, making it a data.frame to spruce it up easier in kable
columnNASums <- data.frame(colSums(is.na(brew_beer)))

#change name of number column
names(columnNASums) <- c("# Beers")

#printing nicely
knitr::kable(columnNASums[which(columnNASums[,1]>0),, drop = FALSE]) %>%
  kable_styling(full_width = F)
```

The IBU measurements are missing account for nearly half of the beers in the data set.

*RESULTS WILL BE SKEWED*



